### 2018-07-10 source/Ch8-hiresEGS/hires-EGS.tex
backup

\chapter{Imaging the Extended Groth Strip with the full LOFAR array}\label{section.EGS.hires}
\minitoc
\section{Aims \& Methodology}

@@

\pg
In other words, we seek to estimate whether, in practice, the EGS can, in part or in full, be placed on the same facet as 3C295 - and with what loss in accuracy to directional gain errors. The first direction-dependent effect we must take into account is decorrelation, also known as smearing (see \cref{section.RIME.TimeDep}). This effect causes signal loss on the longest baselines for sources far from the phase centre. A significant preliminary test is therefore to check that we do in fact see any of the LBCS sources - which we do, as will be demonstrated fully below. However, this is not the end of the problem: decorrelation causes flux from a source to be ``smeared" (hence its other name) around the source. The DDFacet suite can account for this effect by modelling the direction-dependent PSF, i.e. a PSF which is smeared appropriately as a function of position\footnote{A single smeared PSF is calculated per facet, and this smeared PSF is treated as being constant within the facet: this is a good approximation unless one uses truly massive facets.}

@@

\pg
In RIME terms, when performing our direction-independent calibration, we solved for $\hatJones_{3C295,p}^{t,\nu}$ (see \cref{3c295.dirin.cal}), assuming that it was the dominant source of signal in the sky. By applying the inverse of this gain solution to the visibilities, we image the \textit{corrected visibilities}:
\begin{align}
\Vis_{pq}^\mathrm{corr} (t,\nu) &= \hatJones_{3C295,p}^{-1}(t,\nu) \Vis_{pq}(t,\nu) \left(\hatJones_{3C295,q}(t,\nu)\right)^{-H}\\
                        &=  \hatJones_{3C295,p}^{-1}(t,\nu) \left(\sum_s \Jones_{s,p}(t,\nu)\Bmatrix_{pq}(t,\nu) \Jones^H_{s,q}(t,\nu)\right) \hatJones^{-H}_{3C295,q}(t,\nu)\\
                        &= \tildeGjones_{p}(t,\nu) \sum_s \left( \sum_s \tildeEjones_{s,p}(t,\nu)\Bmatrix_{pq}(t,\nu) \tildeEjones_{s,q}^H(t,\nu)\right) \tildeGjones^H_{q}(t,\nu)
\end{align}
where
\begin{align}
\tildeGjones_{p}(t,\nu) &= \Ghatjones_{p}^{-1}(t,\nu)\Gjones_{p}(t,\nu)\\
\tildeEjones_{s,p}(t,\nu) &= \Ehatjones_{3C295,p}^{-1}(t,\nu)\Ejones_{s,p}(t,\nu)
\end{align}                        
Provided that our gain solutions are exactly right (i.e. that $\hatJones_p^{t\nu}=\Jones_p^{t\nu}$ for all $p,t,\nu$) then $\tildeGjones=1 \forall p,t,\nu$. If the calibration problem is sufficiently well-conditioned (i.e. good signal-to-noise, many baselines) and the sky model sufficiently complete, then this is a reasonable assumption to make. However, for all $s\ne 3C295$, $\tildeEjones\ne 1$. This means that, depending on how different the $\Ejones$ terms will be between a given direction and 3C295, direction-dependent calibration artefacts will be introduced in the field.

\pg
In practice, both $\tildeGjones$ and $\tildeEjones$ will contribute to image deterioration: in our case, we only model part of the sky to which our array is sensitive, and so $\tildeGjones\ne 1$ differently for various antennas, times and frequencies. Similarly, we do not perform direction-dependent calibration: as distance increases from 3C295, it is increasingly likely that the ionosphere through which non-3C295 sources' signal propagates before reaching our antennas will vary from the ionosphere through which the signal from 3C295 will vary - and because the ionosphere is turbulent, this behaviour is unpredictable as a function of distance or frequency. The departure of both $\tildeGjones$ and $\tildeEjones$ from unity as a function of direction is called \textit{directional gain errors}, as it is caused by direction-dependent effects.

\pg
Assuming that the ionosphere is relatively smooth and dependent only on distance from 3C295, we attempt to characterise the impact of directional gain errors and their associated image-plane DDEs on various calibrator sources around 3C295 and the EGS.

\pg
We select 8 LBCS\footnote{Long Baseline Calibrator Survey \citepads{2016A&A...595A..86J}} sources, which are all the ``good" sources within 5 degrees of the centre of the EGS. \cref{lbcs-coverage-image} shows the position of these calibrator sources with respect to the EGS. If we compare \cref{lbcs-coverage-image} and \cref{plot.EGS.lofar.widefield}, we can immediately see that the EGS lies in two separate facets of our direction-dependent calibration, which are the eastern neighours of the 3C295 facet. Neither of these EGS facets contains a LBCS source (indeed, if they did, our work would have perhaps been easier); however, nearby facets do contain one or multiple LBCS calibrators. If the effect of directional gain errors for these sources is comparable to those for the EGS facets, then we can estimate their impact on an eventual high-resolution image of the EGS.

@@

\section{Testing Directional Gain Errors - LOBOS Sources in the Primary Beam}\label{section.decorr}

@@

%describe LOBOS catalogue \& sources chosen  - why, how
\begin{table}[h!]
\begin{tabular}{ccccc}
\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
\end{tabular}
\caption{\label{table.LOBOS.sources}Table giving the positions of our 8 LBCS sources and their distance from both the centre of the EGS and 3C295.}
\end{table}
\begin{figure}[h!]
\includegraphics[width=\linewidth]{images/{image_full_ampphase_di_m.NS_Smooth.noise01.fitslbcs.egs.positions}.png}
\caption{Position of the calibrator sources around the EGS, overlaid on the widefield EGS image. LBCS is 3C295.}
\label{lbcs-coverage-image}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs1.overlay.5sigma}.png}}
\caption{\label{fig.lbcs1.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS1 over the widefield image.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs1.overlay.15sigma}.png}}
\caption{\label{fig.lbcs1.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS1 over the widefield image.}
\end{subfigure}
\caption{\label{fig.lbcs1.overlay} Overlays of international LOFAR image of LBCS1 over the widefield image.}
\end{figure}

% % % ADD 2 AND 3 WHEN THEY ARE DONE

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs4.overlay.5sigma}.png}}
\caption{\label{fig.lbcs4.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS4 over the widefield image.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs4.overlay.15sigma}.png}}
\caption{\label{fig.lbcs4.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS4 over the widefield image.}
\end{subfigure}
\caption{\label{fig.lbcs4.overlay} Overlays of international LOFAR image of LBCS4 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs5.overlay.5sigma}.png}}
\caption{\label{fig.lbcs5.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS5 over the widefield image.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs5.overlay.15sigma}.png}}
\caption{\label{fig.lbcs5.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS5 over the widefield image.}
\end{subfigure}
\caption{\label{fig.lbcs5.overlay} Overlays of international LOFAR image of LBCS5 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs7.overlay.5sigma}.png}}
\caption{\label{fig.lbcs7.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS7 over the widefield image.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs7.overlay.15sigma}.png}}
\caption{\label{fig.lbcs7.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS7 over the widefield image.}
\end{subfigure}
\caption{\label{fig.lbcs7.overlay} Overlays of international LOFAR image of LBCS7 over the widefield image.}
\end{figure}

\clearpage
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics[width=0.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs8.overlay.5sigma}.png}}
\caption{\label{fig.lbcs8.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS8 over the widefield image. LBCS8 is too faint to appear at the $15\sigma$ level.}
\end{figure}

\pg
A few things are immediately apparent: LBCS1 seems to be a binary source, and both it and LBCS8 seem to be mostly free from direction-dependent calibration effects. LBCS4, LBCS5, and LBCS7, however, show the presence of very strong artefacts, even at the $15\sigma$ level. In the case of LBCS8, the absence of artefacts is likely due to the fact that the source itself is extremely faint.

\pg
To quantify the effect of directional gain errors, we begin by making the approximation that they are a function of distance from the calibrator source only, rather than specifically a function of direction and distance. This is simply because we don't have enough calibrators to do otherwise. Secondly, we treat the effect of directional gain errors as \textit{increasing the noise} around a source compared to what the thermal noise is expected to be around it: this thermal noise is expected to increase as a function of distance from the phase centre because we apply a beam-correction during imaging (i.e. because we do not image apparent flux but integrated flux). \cref{fig.noiserms.distEGS} shows this effect in practice; note that the noise estimation around LBCS6 (3C295) is biased because we are still quite close to this extremely bright source, and thus likely have a large contribution from artefacts. This explains its outlier behaviour.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/RMSvsDistFromEGS.png}
\caption{Thermal noise near LBCS sources as a function of distance from the phase centre.}
\label{fig.noiserms.distEGS}
\end{figure}

@@

\pg
\textcolor{red}{need LBCS3 to finish imaging to really conclude here, sadly. the EGS lies in a triangle between 3C295, LBCS4 and LBCS3 , so if LBCS3 is fine, then the project is fine!}

\clearpage
\section{Patchwise Imaging of the EGS using LOFAR International Stations}

\pg
\textcolor{red}{check that nancep3 can be used for another week and talk about source subtraction; if project cannot go ahead in the next 2 weeks (which it should be able to if we don't self-cal) then scrap this section and move straight on to conclusions \& future work.}

>>>>

\section{Aims \& Methodology}

@@

\begin{table}[h!]
\begin{tabular}{ccccc}
\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
\end{tabular}
\caption{\label{table.LOBOS.sources}Table giving the positions of our 8 LBCS sources and their distance from both the centre of the EGS and 3C295.}
\end{table}
\begin{figure}[h!]
\includegraphics[width=\linewidth]{images/{image_full_ampphase_di_m.NS_Smooth.noise01.fitslbcs.egs.positions}.png}
\caption{Position of the calibrator sources around the EGS, overlaid on the widefield EGS image. LBCS is 3C295.}
\label{lbcs-coverage-image}
\end{figure}

\pg
In other words, we seek to estimate whether, in practice, the EGS can, in part or in full, be placed on the same facet as 3C295 - and with what loss in accuracy to directional gain errors. The first direction-dependent effect we must take into account is decorrelation, also known as smearing (see \cref{section.RIME.TimeDep}). This effect causes signal loss on the longest baselines for sources far from the phase centre. To do this, we make use of the LOFAR Long-Baseline Calibrator Survey \citepads[LBCS,]{2016A&A...595A..86J}. This is a survey of suitable compact sources to use as calibrators for international LOFAR baselines: these sources are expected to consist mainly of self-absorbed synchrotron radio cores. The survey's source selection procedure is as follows: it selects all sources identified as single sources in WENSS \citepads{1997A&AS..124..259R}. Then, based on the work of \citetads{2015A&A...574A..73M}, which found that the main predictors for compact structure were a high total WENSS flux density and a flat low-frequency spectrum (in line with the expectation above), a selection was made on the basis of spectral index and WENSS flux density for all these sources. %Because the survey aims to span the whole northern sky, there is no guarantee that LBCS sources are single sources

\pg
A significant preliminary test is therefore to check that we do in fact see any of the LBCS sources. However, this is not the end of the problem: decorrelation causes flux from a source to be ``smeared" (hence its other name) around the source. The DDFacet suite can account for this effect by modelling the direction-dependent PSF, i.e. a PSF which is smeared appropriately as a function of position\footnote{A single smeared PSF is calculated per facet, and this smeared PSF is treated as being constant within the facet: this is a good approximation unless one uses truly massive facets.}

@@

\pg
In RIME terms, when performing our direction-independent calibration, we solved for $\hatJones_{3C295,p}^{\nu}$ (see \cref{3c295.dirin.cal}), assuming that it was the dominant source of signal in the sky. By applying the inverse of this gain solution to the visibilities, we image the \textit{corrected visibilities}:
\begin{align}
\Vis_{pq}^\mathrm{corr} (\nu) &= \hatJones_{3C295,p}^{-1}(\nu) \Vis_{pq}(\nu) \left(\hatJones_{3C295,q}(\nu)\right)^{-H}\\
                        &=  \hatJones_{3C295,p}^{-1}(\nu) \left(\sum_s \Jones_{s,p}(\nu)\Bmatrix_{pq}(\nu) \Jones^H_{s,q}(\nu)\right) \hatJones^{-H}_{3C295,q}(\nu)\\
                        &= \tildeGjones_{p}(\nu) \sum_s \left( \sum_s \tildeEjones_{s,p}(\nu)\Bmatrix_{pq}(\nu) \tildeEjones_{s,q}^H(\nu)\right) \tildeGjones^H_{q}(\nu)
\end{align}
where
\begin{align}
\tildeGjones_{p}(\nu) &= \Ghatjones_{p}^{-1}(\nu)\Gjones_{p}(\nu)\\
\tildeEjones_{s,p}(\nu) &= \Ehatjones_{3C295,p}^{-1}(\nu)\Ejones_{s,p}(\nu)
\end{align}                        
where we have muted the time-dependence of all variables for the sake of readability. Provided that our gain solutions are exactly right (i.e. that $\hatJones_p^{\nu}=\Jones_p^{\nu}$ for all $p,\nu$) then $\tildeGjones=\I \forall p,\nu$. If the calibration problem is sufficiently well-conditioned (i.e. good signal-to-noise, many baselines) and the sky model sufficiently complete, then this is a reasonable assumption to make. However, for all $s\ne 3C295$, $\tildeEjones\ne 1$. This means that, depending on how different the $\Ejones$ terms will be between a given direction and 3C295, direction-dependent calibration artefacts will be introduced in the field.

\pg
In practice, both $\tildeGjones$ and $\tildeEjones$ will contribute to image deterioration: in our case, we only model part of the sky to which our array is sensitive, and so $\tildeGjones\ne 1$ differently for various antennas, times and frequencies. Similarly, we do not perform direction-dependent calibration: as distance increases from 3C295, it is increasingly likely that the ionosphere through which non-3C295 sources' signal propagates before reaching our antennas will vary from the ionosphere through which the signal from 3C295 will vary - and because the ionosphere's behaviour is complex, this behaviour is unpredictable as a function of distance or frequency. The departure of both $\tildeGjones$ and $\tildeEjones$ from unity as a function of direction is called \textit{directional gain errors}, as it is caused by direction-dependent effects.

\pg
Assuming that direction-dependent effects (beam, ionosphere) are relatively smooth and dependent only on distance from 3C295, we attempt to characterise the impact of directional gain errors and their associated image-plane DDEs on various calibrator sources around 3C295 and the EGS.

\pg
We select 8 LBCS sources, which are all the ``good" sources within 5 degrees of the centre of the EGS. \cref{lbcs-coverage-image} shows the position of these calibrator sources with respect to the EGS. If we compare \cref{lbcs-coverage-image} and \cref{plot.EGS.lofar.widefield}, we can immediately see that the EGS lies in two separate facets of our direction-dependent calibration, which are the eastern neighours of the 3C295 facet. Neither of these EGS facets contains a LBCS source (indeed, if they did, our work would have perhaps been easier); however, nearby facets do contain one or multiple LBCS calibrators. If the effect of directional gain errors for these sources is comparable to those for the EGS facets, then we can estimate their impact on an eventual high-resolution image of the EGS.

% % %here

@@

\section{Testing Directional Gain Errors - LBCS Sources in the Primary Beam}\label{section.decorr}

@@

%describe LOBOS catalogue \& sources chosen  - why, how

@@

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs1.overlay.5sigma}.png}
\caption{\label{fig.lbcs1.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS1 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs2.overlay.5sigma}.png}
\caption{\label{fig.lbcs2.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS2 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs3.overlay.5sigma}.png}
\caption{\label{fig.lbcs3.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS3 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs4.overlay.5sigma}.png}
\caption{\label{fig.lbcs4.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS4 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs5.overlay.5sigma}.png}
\caption{\label{fig.lbcs5.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS5 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs6.overlay.5sigma}.png}
\caption{\label{fig.lbcs6.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS6 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs7.overlay.5sigma}.png}
\caption{\label{fig.lbcs7.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS7 over the widefield image.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs8.overlay.5sigma}.png}
\caption{\label{fig.lbcs8.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS8 over the widefield image.}
\end{figure}

%
%
%% % % ADD 2 AND 3 WHEN THEY ARE DONE
%
%
%\begin{figure}[h!]
%\centering
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs4.overlay.5sigma}.png}}
%\caption{\label{fig.lbcs4.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS4 over the widefield image.}
%\end{subfigure}
%\hfill
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs4.overlay.15sigma}.png}}
%\caption{\label{fig.lbcs4.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS4 over the widefield image.}
%\end{subfigure}
%\caption{\label{fig.lbcs4.overlay} Overlays of international LOFAR image of LBCS4 over the widefield image.}
%\end{figure}
%
%\begin{figure}[h!]
%\centering
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs5.overlay.5sigma}.png}}
%\caption{\label{fig.lbcs5.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS5 over the widefield image.}
%\end{subfigure}
%\hfill
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs5.overlay.15sigma}.png}}
%\caption{\label{fig.lbcs5.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS5 over the widefield image.}
%\end{subfigure}
%\caption{\label{fig.lbcs5.overlay} Overlays of international LOFAR image of LBCS5 over the widefield image.}
%\end{figure}
%
%\begin{figure}[h!]
%\centering
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs7.overlay.5sigma}.png}}
%\caption{\label{fig.lbcs7.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS7 over the widefield image.}
%\end{subfigure}
%\hfill
%\begin{subfigure}{.8\textwidth}
%\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs7.overlay.15sigma}.png}}
%\caption{\label{fig.lbcs7.overlay.15sigma} $15\sigma$ overlay of international LOFAR image of LBCS7 over the widefield image.}
%\end{subfigure}
%\caption{\label{fig.lbcs7.overlay} Overlays of international LOFAR image of LBCS7 over the widefield image.}
%\end{figure}
%
%\clearpage
%\begin{figure}[h!]
%\centering
%\resizebox{\hsize}{!}{\includegraphics[width=0.8\linewidth]{images/{image_full_ampphase_di_m.NS.int.restored.fits.lbcs8.overlay.5sigma}.png}}
%\caption{\label{fig.lbcs8.overlay.5sigma} $5\sigma$ overlay of international LOFAR image of LBCS8 over the widefield image. LBCS8 is too faint to appear at the $15\sigma$ level.}
%\end{figure}

\pg
A few things are immediately apparent: LBCS1 seems to be a binary source, and LBCS3 seems to suffer strongly from directional gain errors (the source structure in stripes seems unlikely to be physical). LBCS4 seems almost unaffected, which would indicate that the portions of the EGS which lie between 3C295 and LBCS4 are not strongly affected by directional gain errors. The fact that LBCS3 seems to be affected would corroborate the notion that the characteristic scale of DDE impacts is of the order of just over a degree.%, LBCS5, and LBCS7, however, show the presence of very strong artefacts, even at the $15\sigma$ level. In the case of LBCS8, the absence of artefacts is likely due to the fact that the source itself is extremely faint.

\pg
Of course, it is not enough to eyeball this: we therefore proceed to quantify the impact of directional gain errors for different sources in the field. To do this, we begin by making the approximation that they are a function of distance from the calibrator source only, rather than specifically a function of direction and distance. This is simply because we don't have enough calibrators to do otherwise. Secondly, we treat the effect of directional gain errors as \textit{increasing the noise} around a source compared to what the thermal noise is expected to be around it: this thermal noise is expected to increase as a function of distance from the phase centre because we apply a beam-correction during imaging (i.e. because we do not image apparent flux but integrated flux). \cref{fig.noiserms.distEGS} shows this effect in practice; note that the noise estimation around LBCS6 (3C295) is biased because we are still quite close to this extremely bright source, and thus likely have a large contribution from artefacts. This explains its outlier behaviour.
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/RMSvsDistFromEGS.png}
\caption{Thermal noise near LBCS sources as a function of distance from the phase centre.}
\label{fig.noiserms.distEGS}
\end{figure}

@@

\pg
The most encouraging result shown above, however, is for LBCS3 \& LBCS4: both of these sources appear to be only very slightly affected by directional gain errors. Since the EGS lies between these two calibrator sources and 3C295, this is extremely positive. Similarly, the fact that LBCS4 resolves into two separate sources (as does LBCS1, for that matter) would indicate that even neighbouring bright sources do resolve properly - this is not a rigorous test, but it is a fortunate serendipitous observation to be able to make.

%\clearpage
%\section{Patchwise Imaging of the EGS using LOFAR International Stations}
%
%\pg
%\textcolor{red}{check that nancep3 can be used for another week and talk about source subtraction; if project cannot go ahead in the next 2 weeks (which it should be able to if we don't self-cal) then scrap this section and move straight on to conclusions \& future work.}
